# Expert survey results {#sec-app-b}



```{r appdx-b-setup}
#| message: false
#| warning: false
#| cache: false

# test format
is_html <- knitr::is_html_output()
# define paths
response_data_path <- "../../../../data/survey_experts/"
questions_path <- "../questions_EN/"
functions_path <- "../../../functions/"
# source scripts
source("../09_survey_make_plots.R", local = knitr::knit_global())

```


## Additional descriptive results {#sec-app-descr-results}

### Invasion stadium PrIUS against survey

@fig-stadium displays the invasion stadium as assigned in the PrIUS report [@dhondt_een_2022] against the invasion stadium as reported in the expert survey.
Species for which discrepancies occur are highlighted.

```{r plot-stadium}
#| label: "fig-stadium"
#| fig-cap: "Invasion stadium according to PrIUS and the expert survey"
#| fig-width: 10
#| fig-height: 8
#| out-width: 90%

plot_stadium

```

{{< pagebreak >}}

### Response proportions "unknown / I do not know" 

@fig-score-props-1 shows the proportion of responses that fall into the category "unknown / I do not know" across all questions per species.
As in @fig-ranking, species are sorted according to their overall urgency scores nested within their overall feasibility scores.


```{r plot-score-props-1}
#| label: "fig-score-props-1"
#| fig-cap: "Relative proportions of responses 'unknown / I do not know' per species"
#| fig-width: 10
#| fig-height: 16
#| out-width: 90%

plot_unknown_prop_species

```

{{< pagebreak >}}

@fig-score-props-2 shows the proportion of responses that fall into the category "unknown / I do not know" across all species per question.

```{r plot-score-props-2}
#| label: "fig-score-props-2"
#| fig-cap: "Relative proportions of responses 'unknown / I do not know' per question"
#| fig-width: 10
#| fig-height: 10
#| out-width: 90%

plot_unknown_prop_question

```

{{< pagebreak >}}

@fig-score-patterns, finally, shows the proportion of responses that fall into the category "unknown / I do not know" per question and per species.
Species are sorted according to their overall urgency scores nested within their overall feasibility scores.

```{r plot-score-patterns}
#| label: "fig-score-patterns"
#| fig-cap: "Pattern of responses 'unknown / I do not know' per species and question"
#| fig-width: 10
#| fig-height: 16
#| out-width: 90%

plot_unknown_pattern

```

{{< pagebreak >}}

### Monitoring methods

@fig-method-freq displays the monitoring methods reported in the survey sorted by frequency.
Note that per species multiple methods could be reported. 
Methods are grouped into categories distinguished by color.

```{r plot-method-freq}
#| label: "fig-method-freq"
#| fig-cap: "Absolute frequencies of monitoring methods reported"
#| fig-width: 10
#| fig-height: 12
#| out-width: 90%

plot_meth_freq

```

{{< pagebreak >}}

@fig-method-pairs shows all pairwise combinations of monitoring methods reported in the survey sorted by frequency. 

```{r plot-method-pairs}
#| label: "fig-method-pairs"
#| fig-cap: "Absolute frequencies of pairwise combinations of monitoring methods reported"
#| fig-width: 10
#| fig-height: 14
#| out-width: 90%

plot_meth_netw

```


{{< pagebreak >}}

### Existing monitoring efforts

@fig-moni-pattern provides an overview over existing monitoring efforts reported per species. 
Monitoring efforts are sorted by the number of species they cover (highest coverage left).
Species are grouped by kingdom, and within groups sorted according to their overall urgency scores nested within their overall feasibility scores.


```{r plot-moni-pattern}
#| label: "fig-moni-pattern"
#| fig-cap: "Existing monitoring efforts reported per species"
#| fig-width: 10
#| fig-height: 16
#| out-width: 90%

plot_moni_patt

```

