# Results expert survey

In this section we provide the results of the expert survey.
We briefly describe the response rate and proceed with the results of the prioritization process.
We present results in terms of 3 scenarios / species focus groups...
Results for the full set of species we have collected data on can be found in appendix X.
This concerns mainly the prioritization results but also a number of descriptive results (?).


```{r survey results setup}
#| message: false
#| warning: false
#| cache: false

# test format
is_html <- knitr::is_html_output()

# define paths
response_data_path <- "../../../../data/survey_experts/"
questions_path <- "../questions_EN/"
functions_path <- "../../../functions/"

# load response & question data
res_comb_upd <- get(load(paste0(response_data_path, "results_combined_upd.rda")))

# source scripts
source("../09_format_priority_tables.R", local = knitr::knit_global())

options(knitr.kable.NA = '')

```

## Response rate

Questionnaire completed for `r res_comb_upd$species |> unique() |> length()` species.
Among these are `r res_comb_upd |> dplyr::filter(on_unionlist) |> dplyr::pull(species) |> unique() |> length()` IAS of union concern 
and `r res_comb_upd |> dplyr::filter(!on_unionlist) |> dplyr::pull(species) |> unique() |> length()` candidate species.


## ANB priority species

### Prioritization

#### Plants

[description]

```{r table prior plants anb}
#| label: tbl-priorplantsanb
#| tbl-cap: Prioritized species-scope combinations for ANB priorty plants

table_filtered_anb_display_list$plant

```



#### Animals

[description]

```{r table prior animals anb}
#| label: tbl-prioranimalsanb
#| tbl-cap: Prioritized species-scope combinations for ANB priorty animals

table_filtered_anb_display_list$animal

```



### Method-based synergies

#### Plants

[description]

```{r table-syn-plants-anb-1}
#| label: tbl-synplantsanb1
#| tbl-cap: Method-based synergies for ANB priorty plants under scope abundance after management

table_syn_anb_plants_display_list$abundance_management

```



#### Animals

[description]

```{r table-syn-animals-anb-1}
#| label: tbl-synanimalsanb1
#| tbl-cap: Method-based synergies for ANB priorty animals under scope distribution after management

table_syn_anb_animals_display_list$distribution_management


```



```{r table-syn-animals-anb-2}
#| label: tbl-synanimalsanb2
#| tbl-cap: Method-based synergies for ANB priorty animals under scope abundance after management


table_syn_anb_animals_display_list$abundance_management

```



## Focus on scope detection

### Prioritization

#### Plants

[description]

```{r table prior plants det}
#| label: tbl-priorplantsdet
#| tbl-cap: Prioritized species-scope combinations for plants among species for which detection is relevant
#| 
table_filtered_det_display_list$plant

```


#### Animals

[description]

```{r table prior animals det}
#| label: tbl-prioranimalsdet
#| tbl-cap: Prioritized species-scope combinations for animals among species for which detection is relevant

table_filtered_det_display_list$animal

```


### Method-based synergies


#### Plants

[description]

```{r table-syn-plants-det-1}
#| label: tbl-synplantsdet1
#| tbl-cap: Method-based synergies for plants under scope detection 

table_syn_det_plants_display_list$detection

```


#### Animals

[description]

```{r table-syn-animals-det-1}
#| label: tbl-synanimalsdet1
#| tbl-cap: Method-based synergies for animals under scope detection 

table_syn_det_animals_display_list$detection


```



## Focus on scope distribution

### Prioritization

#### Plants

[description]

```{r table prior plants dist}
#| label: tbl-priorplantsdist
#| tbl-cap: Prioritized species-scope combinations for plants among species for which distribution is relevant
#| 
table_filtered_dist_display_list$plant

```



#### Animals

[description]


```{r table prior animals dist}
#| label: tbl-prioranimalsdist
#| tbl-cap: Prioritized species-scope combinations for animals among species for which distribution is relevant

table_filtered_dist_display_list$animal

```


### Method-based synergies

#### Plants

[description]


```{r table-syn-plants-dist-1}
#| label: tbl-synplantsdist1
#| tbl-cap: Method-based synergies for plants under scope distribution 

table_syn_dist_plants_display_list$distribution

```



#### Animals

[description]

```{r table-syn-animals-dist-1}
#| label: tbl-synanimalsdist1
#| tbl-cap: Method-based synergies for animals under scope distribution 

table_syn_dist_animals_display_list$distribution

```


