---
# generic options. See ?INBOmd::pdf_report() for full list and more details.
title:
# subtitle:
author: 
  - name:
      given: Firstname1
      family: Last name1
    email: "firstname1.lastname1@inbo.be"
    orcid: 0000-0002-1825-0097
    affiliation: Research Institute for Nature and Forest (INBO)
    corresponding: true
  - name:
      given: Firstname2
      family: Lastname2
    email: "firstname2.lastname2@inbo.be"
    orcid: 0000-0000-0000-0000
    affiliation: Research Institute for Nature and Forest (INBO)
style: Flanders
lang: en
# report specific options. See ?INBOmd::pdf_report() for full list and more details.
reviewer:
  - name:
      given: Firstname
      family: Lastname
    email: "firstname.lastname@inbo.be"
    orcid: 0000-0000-0000-0000
    affiliation: Research Institute for Nature and Forest (INBO)
funder: Research Institute for Nature and Forest (INBO)
rightsholder: Research Institute for Nature and Forest (INBO)
community: inbo
keywords: keyword1
# year:
cover_photo:
# cover_description:
# cover:
# doi:
# reportnr:
# ordernr:
# depotnr: 
client:
  - Organisation (remove client section when not needed)
  - Street number
  - Postal code city
  - https://www.website.com
# client_logo: uncomment and replace with path to logo or remove line
# pdf report specific options. See ?INBOmd::pdf_report() for full list and more details.
embargo:
print:
  copies: 0
  motivation:
  pages:
bibliography: references.bib
link-citations: TRUE
site: bookdown::bookdown_site
output:
  # Don't run the format below.
  # Only required for RStudio to recognise the project as a bookdown project.
  # Hence don't use "Build all formats".
  INBOmd::pdf_report: default
  INBOmd::gitbook: default
  INBOmd::epub_book: default
  bookdown::dontrun: default
---

# Acknowledgements {-}

Text here.


```{r setup, include = FALSE}
library(knitr)
opts_chunk$set(
  echo = FALSE,
  eval = TRUE,
  cache = FALSE,
  warning = TRUE,
  error = FALSE,
  message = TRUE
)
library(tidyverse)
library(INBOtheme)
if (interactive()) {
  theme_set(theme_inbo(base_size = 10))
} else {
  switch(
    opts_knit$get("rmarkdown.pandoc.to"),
    html = {
      opts_chunk$set(dev = "png", dpi = 72)
      theme_set(theme_inbo(base_size = 12))
    },
    latex = {
      opts_chunk$set(dev = "cairo_pdf", dpi = 300)
      theme_set(theme_inbo(base_size = 9))
      update_geom_defaults("point", list(size = 1.5))
    },
    epub3 = {
      opts_chunk$set(dev = "png", dpi = 300)
      theme_set(theme_inbo(base_size = 12))
    }
  )
}
```

<!--chapter:end:index.Rmd-->

# Abstract {-}

\benglish

Summary here.

\eenglish

\newpage

<!--chapter:end:00_1_abstract.Rmd-->

# Dutch abstract {-}

Samenvatting hier.

\newpage

<!--chapter:end:00_2_abstract_dutch.Rmd-->

# Dutch executive summary {-}

Uitgebreide samenvatting hier.

\newpage

<!--chapter:end:00_3_executive_summary_dutch.Rmd-->

<!-- hier komt de inhoudstafel bij pdf -->
\mainmatter

# Introduction

\benglish

Text here.

\eenglish

\newpage

<!--chapter:end:01_introduction.Rmd-->

# Demand side

## Client

\benglish

Text here.

\eenglish

\newpage

<!--chapter:end:02_1_1_demand_side_client.Rmd-->

<!--# Demand side -->
## Users

\benglish

Text here.

\eenglish

\newpage

<!--chapter:end:02_1_2_demand_side_users.Rmd-->

# Scientific basis

\benglish

Text here.

\eenglish

\newpage

<!--chapter:end:02_2_scientific_basis.Rmd-->

# Existing monitoring

\benglish

Text here.

\eenglish

\newpage

<!--chapter:end:02_3_existing_monitoring.Rmd-->

# Conditions

\benglish

Text here.

\eenglish

\newpage

<!--chapter:end:02_4_conditions.Rmd-->

```{r references, results = "asis", echo = FALSE}
# insert the references at this position
# set appendix = FALSE, when the report has no appendix
INBOmd::references(appendix = TRUE)
```

<!--chapter:end:zz_01_references.Rmd-->

# First appendix

\benglish

Text here.

\eenglish

\newpage

<!--chapter:end:zz_02_appendix.Rmd-->

# Glossary

\benglish

Foo

:   description of foo

Bar

:   descirption of bar

\eenglish

\newpage

<!--chapter:end:zz_03_glossary.Rmd-->

