# Outtakes

## Scenario 1: ANB priority species (plants)

- Method synergies for plants & scope abundance

```{r scenario-anb-priorities-syn-plants-abd}
#| results: asis

table_syn_anb_plants_display_list$abundance |> unclass() |> cat()

```

## Scenario 1: ANB priority species (plants)

- Method synergies for plants & scope abundance after management

```{r scenario-anb-priorities-syn-plants-abd-after}
#| results: asis

table_syn_anb_plants_display_list$abundance_management |> unclass() |> cat()

```


## Scenario 1: ANB priority species (animals)

- Method synergies for animals & scope abundance after management

```{r scenario-anb-priorities-syn-animals-abd-after}
#| results: asis

table_syn_anb_animals_display_list$abundance_management |> unclass() |> cat()

```


## Scenario 2: Focus on detection (plants)

- Method synergies for plants & scope abundance

```{r scenario-det-priorities-syn-plants-abd}
#| results: asis

table_syn_det_plants_display_list$abundance |> unclass() |> cat()

```

## Scenario 2: Focus on detection (plants)

- Method synergies for plants & scope distribution after management

```{r scenario-det-priorities-syn-plants-dist-after}
#| results: asis

table_syn_det_plants_display_list$distribution_management |> unclass() |> cat()

```


## Scenario 3: Focus on distribution (plants)

- Method synergies under scope abundance

```{r scenario-dist-priorities-syn-plants-abd}
#| results: asis

table_syn_dist_plants_display_list$abundance |> unclass() |> cat()

```


## Scenario 3: Focus on distribution (plants)

- Method synergies under scope distribution after management

```{r scenario-dist-priorities-syn-plants-dist-after}
#| results: asis

table_syn_dist_plants_display_list$distribution_management |> unclass() |> cat()

```
