# Prioritization

## Approach

- Question:
  - Which surveillance/monitoring scopes are *relevant*/*feasible* and have *high priority* for which species?
- Propose a range of scopes
  - detection, distribution, abundance etc.
- Determine relevance/feasibility & priority based on survey results
- Present result scenarios to facilitate decision making

::: {.notes}
- aanpak in grote lijnen
- vertrokken van deze vraag
- om die te kunnen beantwoorden moesten we eerst een aantal scopes voorstellen
  - deze deels gebaseerd op de EU regulatie en deels op typische meetnet toepassingen 
  - in short: ..., straks meer informatie
  - hier wel input van de opdrachgever nodig, dus we willen hier graag jullie feedback

:::

## Surveillance/monitoring scopes

- Detection
- Inventory
- (Change in) Distribution
- (Change in) Abundance 
- (Change in) Distribution after managenment
- (Change in) Abundance after management

::: {.notes}
- Wat zouden concrete invullingen van deze scopes zijn?

notas: 

- Ik verander in het vervolg de volgorde en begin met de (voor mij) duidelijke scopes
  - De invullingen van de latere bouwen ook deels verder op die van de vroegere
- Dit zou ik hier zeker ook vermelden om verwarring te vorkomen
- Het is natuurlijk ook mogelijk om de volgorde op deze overzichtsslide aan te passen
  - (enkel in de tabellen later zou ik dat niet doen)


:::


## Scope distribution

- *status distribution area* 
  - size of area
  - presence and absence in Flanders or potential distribution area
    - distribution area deduced from presence/absence in 10x10 km grid cells
  - reported every X years, with X $\leq$ 6
- *trend distribution area*
  - changes (relative size and direction) in size of area
  - meaningful/critical changes to be defined 


::: {.notes}
- potential distribution area: known distribution area plus buffer
- buffer depends on the presence of potential habitats neighboring the area of distribution, mobility of the species
- 10 x 10 km grid cells of e.g. the EEA reference grid (~185 across Flanders)
- alternatives: 5 x 5 km (~740 across Flanders); 1 x 1 km (~1850 across Flanders)
- EU: "including information regarding migratory or reproductive patterns"?
- possibility to complement ad hoc data with structured observations in grid cells where the species has not been found present in ad hoc data
- distribution area needs to be known (e.g., previously reported), if that's not the case ...
:::

## Scope inventory

- *status of distribution area*
-   determine location of (potential) distribution area
    - candidate areas via list of all possible habitats (e.g., expert judgement)
    - presence and absence in candidate areas
- analogies with scope distribution


::: {.notes}
- status w.r.t. location of area
- analogies with distribution surveillance
  - area divided into grid cells/resolution in terms of grid cells
  - certain search effort required

:::

## Scope abundance

- *status abundance*
  - population size (relative or absolute)
  - in a distribution or other selected area(s)
- *trend abundance*
  - changes (size and direction) in population size


::: {.notes}
- population is defined as the set of individuals of a species present in a biogeographical region within Flanders
- selected areas: e.g., invasion hotspot, area of management

:::

## Scope distribution after management

- status and/or trend in size of area
  - cf. scope distribution
- 'after management'
  - evaluating temporal and spatial correlates
- *Eradication*
  - absence in confined area of management with confined follow-up period
  - limited in space and time
- *Containment*
  - status and/or trend of size of distribution area within confined area of management
  - limited in space, but not time 

::: {.notes}
- 'after management':
  - continuum from determining a causal effect of a measure to evaluating temporal and spatial correlates (assuming some causal effect of a measure)
  - we will probably rather be close to the latter end than the former
- containment opvolgen: mogelijk op langere termijn

:::



## Scope abundance after management

- status and/or trend in size of population
  - cf. scope abundance
- 'after management'
  - cf. scope distribution after management
- *Eradication*
  - cf. distribution after management
- *Population control*
  - status and/or trend of size of population within confined area of management
  - limited in space, but not time


## Scope detection

- presence in defined area
  - around known (potential) places of introduction and sites of sporadic occurrence
  - possibly also in/around conservation area
- analogies with scope distribution


::: {.notes}
- in/around conservation area: 
  - in conservation area
  - periphery of conservation area
  - buffer area around conservation area
- analogies with scope distribution
  - area divided into grid cells/resolution in terms of grid cells
  - certain search effort required

:::


::: {.notes}
- Feedback re scopes?
  - Wilt de stuurgroep aanpassingen/zijn er suggesties?

:::


## Species-scope table illustration

```{r species-scope-table}
#| results: asis

table_base_illu_list_upd$stadium |> 
  dplyr::mutate(
    scope_verbose = NA_character_
  )|>
  dplyr::arrange(species) |>
  make_table_display(
    data_table = _,
    cols_id = c("species"),
    cols_addon = NULL,
    footnote = NULL) |> 
  unclass() |> cat()

```



## Exclusion criteria

- Exclude species-scope combinations that are irrelevant or infeasible
  - Invasion stadium
  - Distribution area known (exclusion inventory)
  - Species not managed
  - Information (presence-absence/abundance) not relevant to evaluate management
  - No method to measure abundance
- Exclude species-scope combinations with low priority
  - Opportunistic data are considered representative
  - Introduction sites/distribution area unknown
  - Global feasibility & urgency scores reach critical value
 

::: {.notes}
- we gebruiken bepaalde individuele survey vragen om uit te sluiten
- maar ook de overal survey score (drempel-waarde mediaan)

nota's: 

- hier misschien een diagram/visualisatie ipv. tekst?

:::


## Species-scope table illustration

```{r species-scope-table-filled}
#| results: asis

table_filtered_illu_final_display |> unclass() |> cat()

```



::: {.notes}
nota's: 

- hier genoeg tijd voor uitleg nemen

:::



## Species-scope table

- [show full tables in html document]
- [include link here]

