## Scenario 1: ANB priority species {.smaller}


::::: columns
::: {.column width="50%"}
High priority in survey

```{r scenario-anb-priorities-species-high}
#| results: "asis"

make_list_species_display <- function(.species){
  .species_tmp <- .species |> 
  dplyr::arrange(kingdom |> dplyr::desc(), vern_name_nld) |>
  dplyr::mutate(vern_name_nld_col = kableExtra::cell_spec(
    x = vern_name_nld,
    color = dplyr::case_when(
      grepl("plant", kingdom) ~ col_plants,
      grepl("dier", kingdom) ~ col_animals
    )
  )) |>
  dplyr::select(vern_name_nld_col, kingdom) |>
  dplyr::mutate(id = dplyr::row_number(), .by = kingdom) |>
  tidyr::pivot_wider(
    id_cols = id,
    names_from = kingdom, values_from = vern_name_nld_col, 
    values_fill = NA
    ) |>
  dplyr::select(-id)
.species_tmp |> 
  knitr::kable(
    format = "html",
    escape = FALSE,
    col.names = c("plant", "animal"),
    table.attr = 'data-quarto-disable-processing="true"'
  ) |> kableExtra::kable_styling(
      bootstrap_options = c("condensed", "hover"), 
      full_width = FALSE,
      position = "left",
      font_size = .font_size
    ) |>
  kableExtra::row_spec(
    row = .species_tmp |> nrow() |> seq_len(),
    extra_css = "border-bottom: 0px"
  )
}

species_anb_highprior |> make_list_species_display()

```
:::

::: {.column width="50%"}
Low priority in survey

```{r scenario-anb-priorities-species-low}
#| results: "asis"

species_anb_lowprior |> make_list_species_display()

```
:::
:::::

::: notes
- voor de rest geen data
- reden low priority in survey: zie html document
:::

## Scenario 1: ANB priority species (plants)

-   Method synergies under scopes detection

```{r scenario-anb-priorities-syn-plants-det}
#| results: asis
#| tab.cap: "Method synergies under scopes detection"

table_syn_anb_plants_display_list$detection |> unclass() |> cat()

```

::: notes

- geen synergien 
- moeraslantaarn enige ANB-plant voor die detectie relevant (en we data hebben)

:::

## Scenario 1: ANB priority species (plants)

-   Method synergies under scopes distribution, abundance, abundance under management

```{r scenario-anb-priorities-syn-plants-dist}
#| results: asis

table_syn_anb_plants_display_list$distribution |> unclass() |> cat()

```

::: notes
- voor scopes distribution, abundance, abundance under management zien we altijd dezelfde groep planten 
  - al die scopes dus relevant
- groep van 5 zoetwater planten (deels terrestrisch)
- synergie: 
  - visuele surveys voor alle 5
  - eDNA: water voor 4 (mogelijk 5)

::: 


## Scenario 1: ANB priority species (plants)

- Clear group:
  - 4 - 5 (partly) freshwater plants
  - Scopes distribution, abundance, and abundance under management are all relevant
    - Focus on management-related scopes most sensible 
  - Aquatic eDNA, visual surveys
- We suggest to analyse for these plants as a group how the scope abundance under management could be filled in



## Scenario 1: ANB priority species (animals)

-   Method synergies under scope detection

```{r scenario-anb-priorities-syn-animals-det}
#| results: asis

table_syn_anb_animals_display_list$detection |> unclass() |> cat()

```

::: notes
- mogelijke synergiën tussen 1 prioritaire en of 1 secundaire of 1 tertiaire soort
  - muskusrat & rivierkreeft via eDNA meest realistisch
  - we weerhouden dus de synergie met de rivierkreeft, maar niet met de eekhoorn op basis van te verschillend habitat

::: 



## Scenario 1: ANB priority species (animals)

-   Method synergies under scope inventory

```{r scenario-anb-priorities-syn-animals-inv}
#| results: asis

table_syn_anb_animals_display_list$inventory |> unclass() |> cat()

```

::: notes
- synergie rode amerikaanse rivierkreeft met tijgermug weinig realistisch wegens verschil in habitat 
- bovendien zouden we beide soorten buiten beschouwing laten 
  - tijgermug: Vlaams en nationaal plan, waaraan gewerkt wordt
  - rivierkreeft: craywatch is bedoeld als inhaalslag en loop nog

::: 

## Scenario 1: ANB priority species (animals) {.scrollable}

-   Method synergies under scope distribution

```{r scenario-anb-priorities-syn-animals-dist}
#| results: asis

table_syn_anb_animals_display_list$distribution |> unclass() |> cat()

```

::: notes

- synergien tussen 3 prioritaire zoetwater soorten en gevlekte rivierkreeft als secundaire soort
  - via eDNA
- synergien tussen 2 terrestrische soorten en tertiaire soort wasbeer
  - cameravallen of visuele surveys

:::

## Scenario 1: ANB priority species (animals)

-   Method synergies under scopes abundance, abundance under management

```{r scenario-anb-priorities-syn-animals-abd}
#| results: asis

table_syn_anb_animals_display_list$abundance |> unclass() |> cat()

```


::: notes
- dezelfde soorten voor abundance en abundance under management
- vooral synergie tussen muntjak en wasbeer realistisch (?)
  - visuele surveys, cameravallen

:::


## Scenario 1: ANB priority species (animals)

-   Method synergies under scope distribution under management

```{r scenario-anb-priorities-syn-animals-dist-after}
#| results: asis

table_syn_anb_animals_display_list$distribution_management |> unclass() |> cat()

```

::: notes
- mogelijke synergie tussen beverrat en wasbeer, maar weinig realistisch wegens verschil in habitat

:::


## Scenario 1: ANB priority species (animals)

-   Larger groups for scope distribution
    -   Group 1
        -   2 terrestrial mammals (Aziatische grondeekhoorn, Chinese muntjak)
        -   Visual surveys, camera traps
        -   1 tertiary mammal (Wasbeer) as potential free rider
    -   Group 2
        -   3 (partly) freshwater species (Afrikaanse klauwkikker, Beverrat, Muskusrat)
        -   Aquatic eDNA
        -   1 secondary crayfish as free rider

::: notes

- Maar wel focus op beheer en we zien deze goepen dan nog eens terugkomen onder scenario 3,
  - waar ze hun kans krijgen

:::

## Scenario 1: ANB priority species (animals)

-   Smaller groups for remaining scopes
    -   Scopes abundance, abundance under management, e.g.,
        -   Chinese muntjak & Wasbeer (free rider) via visual surveys, camera traps
        -   Chinese muntjak & Afrikaanse klauwkikker via visual surveys, passive acoustic monitoring
    -   Scope distribution under management
        -   Beverrat & Wasbeer (free rider) via visual surveys, camera traps
- We suggest to analyse for Chinese muntjak & Afrikaanse klauwkikker how the scope abundance under management could be filled in

::: notes

- Of eerder Chinese muntjak & Wasbeer?

:::
