## Scenario 1: ANB priority species {.smaller}


:::: {.columns}

::: {.column width="50%"}

- High priority in survey

```{r scenario-anb-priorities-species-high}
#| results: "asis"

species_anb |> 
  dplyr::mutate(vern_name_nld_col = kableExtra::cell_spec(
    x = vern_name_nld,
    color = dplyr::case_when(
      grepl("plant", kingdom) ~ col_plants,
      grepl("dier", kingdom) ~ col_animals
    )
  )) |>
  dplyr::filter(grepl("highprior", scope_prior)) |>
  dplyr::arrange(vern_name_nld) |>
  dplyr::pull(vern_name_nld_col) |>
  paste(x = "  -", y = _) |>
  cat(x = _, sep = "\n")

```

:::

::: {.column width="50%"}

- Low priority in survey

```{r scenario-anb-priorities-species-low}
#| results: "asis"

species_anb |> 
    dplyr::mutate(vern_name_nld_col = kableExtra::cell_spec(
    x = vern_name_nld,
    color = dplyr::case_when(
      grepl("plant", kingdom) ~ col_plants,
      grepl("dier", kingdom) ~ col_animals
    )
  )) |>
  dplyr::filter(grepl("lowprior", scope_prior)) |>
  dplyr::arrange(vern_name_nld) |>
  dplyr::pull(vern_name_nld_col) |>
  paste(x = "  -", y = _) |>
  cat(x = _, sep = "\n")

```

:::

::::

::: {.notes}

nota's: 

- hier evtl. 'filtered table' voor anb soorten tonen (zie html document 'plan of action')
- zo duidelijk welke soorten er waarom uitvallen
- tekst kleiner

:::



## Scenario 1: ANB priority species (plants)

- Method synergies under scopes detection

```{r scenario-anb-priorities-syn-plants-det}
#| results: asis
#| tab.cap: "Method synergies under scopes detection"

table_syn_anb_plants_display_list$detection |> unclass() |> cat()

```

::: {.notes}

nota's: 

- tekst als caption aan tabel toevoegen
- ik heb de latijnse namen hier verwijders om plaats te besparen
- de column headers zijn als 'left-aligned' gedefineerd, ik weet niet waarom ze hier naar rechts verschuiven
  - indien nog tijd probeer ik dit nog op te lossen
- "Number of species covered" in presentatie-mode verwijderen en mondeling toelichten

:::




## Scenario 1: ANB priority species (plants)

- Method synergies under scopes distribution, abundance, abundance after management

```{r scenario-anb-priorities-syn-plants-dist}
#| results: asis

table_syn_anb_plants_display_list$distribution |> unclass() |> cat()

```


## Scenario 1: ANB priority species (plants) 

- Clear group: 
  - 4 to 5 (partly) freshwater plants
  - Scopes distribution, abundance, and abundance after management
  - Aquatic eDNA, visual surveys
- We suggest to analyse for these plants as a group which scopes can be filled in how, at which costs

::: {.notes}

nota's: 

- hier graag jullie input
- de suggesties voor de verschillende scenarios nog eens vergelijkend tonen
  - op slide?
  - flipchart? -> misschien handiger

:::


## Scenario 1: ANB priority species (animals)

- Method synergies under scope detection

```{r scenario-anb-priorities-syn-animals-det}
#| results: asis

table_syn_anb_animals_display_list$detection |> unclass() |> cat()

```

## Scenario 1: ANB priority species (animals)

- Method synergies under scope inventory

```{r scenario-anb-priorities-syn-animals-inv}
#| results: asis

table_syn_anb_animals_display_list$inventory |> unclass() |> cat()

```



## Scenario 1: ANB priority species (animals)

- Method synergies under scope distribution

```{r scenario-anb-priorities-syn-animals-dist}
#| results: asis

table_syn_anb_animals_display_list$distribution |> unclass() |> cat()

```

::: {.notes}

nota's: 

- tekst als table caption -> tabel volledig zichtbaar (wasbeer)

:::



## Scenario 1: ANB priority species (animals)

- Method synergies under scopes abundance, abundance after management

```{r scenario-anb-priorities-syn-animals-abd}
#| results: asis

table_syn_anb_animals_display_list$abundance |> unclass() |> cat()

```



## Scenario 1: ANB priority species (animals)

- Method synergies under scope distribution after management

```{r scenario-anb-priorities-syn-animals-dist-after}
#| results: asis

table_syn_anb_animals_display_list$distribution_management |> unclass() |> cat()

```





## Scenario 1: ANB priority species (animals)

- Larger groups for scope distribution
  - Group 1
    - 4 (partly) terrestrial mammals (Nutria, Muskrat, Siberian chipmunk, Chinese muntjac)
    - Visual surveys, camera traps
    - 1 secondary mammal (Racoon) as potential free rider
  - Group 2
    - 3 (partly) freshwater species (African clawed frog, Nutria, Muskrat)
    - Aquatic eDNA
    - 1 secondary crayfish as free rider
    - Heterogenous group, unclear to what extent synergies will play out
  - We suggest to prioritize group 1

::: {.notes}

nota's: 

- hier graag jullie input

:::

## Scenario 1: ANB priority species (animals)

- Smaller groups for remaining scopes
  - Scopes abundance, abundance after management, e.g.,
    - Chinese muntjac & Racoon (free rider) via visual surveys, camera traps
    - Chinese muntjac & African clawed frog via visual surveys, passive acoustic monitoring
  - Scope distribution after management
    - Nutria & Racoon (free rider) via visual surveys, camera traps
- Chinese muntjac & African clawed frog frequently involved

::: {.notes}

nota's: 

- hier graag jullie input

:::


## Scenario 1: ANB priority species (animals)

- Suggestion for the animals in total
  - Analyse for Chinese muntjac which scopes (distribution, abundance, abundance after management) can be filled in how, at which costs
    - For scope distribution also consider Nutria, Muskrat, Siberian chipmunk in addition to Chinese muntjac
  - Analyse for African clawed frog which scopes (distribution, abundance, abundance after management) can be filled in how, at which costs

::: {.notes}

nota's: 

- hier graag jullie input
- alternatief: start with scope distribution and group?
:::
