## Scenario 2: Focus on detection {.smaller} 


:::: {.columns}

::: {.column width="100%"}


- High priority species for which detection is (also) relevant


```{r scenario-det-species}
#| results: "asis"

species_det |> 
  dplyr::mutate(vern_name_nld_col = kableExtra::cell_spec(
    x = vern_name_nld,
    color = dplyr::case_when(
      grepl("plant", kingdom) ~ col_plants,
      grepl("dier", kingdom) ~ col_animals
    )
  )) |>
  dplyr::arrange(vern_name_nld) |>
  dplyr::pull(vern_name_nld_col) |>
  paste(x = "  -", y = _) |>
  cat(x = _, sep = "\n")

```

:::

::::

## Scenario 2: Focus on detection (plants)

- Method synergies under scope detection

```{r scenario-det-priorities-syn-plants-det}
#| results: asis

table_syn_det_plants_display_list$detection |> unclass() |> cat()

```


## Scenario 2: Focus on detection (plants)

- Method synergies under scopes distribution, abundance, distribution after management

```{r scenario-det-priorities-syn-plants-dist}
#| results: asis

table_syn_det_plants_display_list$distribution |> unclass() |> cat()

```


## Scenario 2: Focus on detection (plants) 

- Again clear group: 
  - 3 (partly) terrestrial plants
  - Scope detection
  - Visual surveys
- We suggest to analyse for these plants as a group how and at which costs scope detection can be filled in

::: {.notes}

nota's: 

- hier graag jullie input

:::



## Scenario 2: Focus on detection (animals)

- Method synergies under scope detection

```{r scenario-det-priorities-syn-animals-det}
#| results: asis

table_syn_det_animals_display_list$detection |> unclass() |> cat()

```



## Scenario 2: Focus on detection (animals)

- Method synergies under scope distribution

```{r scenario-det-priorities-syn-animals-dist}
#| results: asis

table_syn_det_animals_display_list$distribution |> unclass() |> cat()

```

## Scenario 2: Focus on detection (animals)

- Method synergies under scope abundance

```{r scenario-det-priorities-syn-animals-abd}
#| results: asis

table_syn_det_animals_display_list$abundance |> unclass() |> cat()

```


## Scenario 2: Focus on detection (animals)

- Larger groups for scope detection
  - Group 1
    - 4 (partly) freshwater species (3 fish, Muskrat)
    - Aquatic eDNA, and for the fish electric fishing, trapping & netting
    - Heterogeneous group, unclear to what extent synergies will play out
  - Group 2
    - 4 (partly) terrestrial species (Muskrat, Small asian mongoose, Chital, House crow)
    - Camera traps, and partly visual surveys
  - We suggest to analyse for both groups how and at which costs scope detection can be filled in

::: {.notes}

nota's: 

- hier graag jullie input
- indische mangoeste via aquatic eDNA?
    
:::


