## Scenario 2: Focus on detection

- Focus on species for which (also) scope detection is relevant
  - Absent & sporadically present species
  - Widespread species absent from conservation areas (prius-stadium 'REL-AAN-GEV-VER-BUI')

- Motivation
  - Importance of intervening in early vs. late invasion stage
  - EU IAS regulation highlights early detection & rapid response ('surveillance')
    - But: responsibility of federal government for detection


::: {.notes}
- INBO: enkel detectie indien soort in natuur opduikt
- introductieplaatsen moeten gekend zijn en voldoende specifiek en lokaal (ook al als filter criterium gebruikt)
- detectie in natuurgebieden moet van belang zijn

:::


## Scenario 2: Focus on detection {.smaller} 

::::: columns
::: {.column width="50%"}

High priority in survey

```{r scenario-det-species}
#| results: "asis"

species_det |> 
  dplyr::filter(grepl("highprior", scope_prior)) |> 
  make_list_species_display()

```
:::
:::::


## Scenario 2: Focus on detection (plants)

- Method synergies under scope detection

```{r scenario-det-priorities-syn-plants-det}
#| results: asis

table_syn_det_plants_display_list$detection |> unclass() |> cat()

```

::: notes

- Mogelijke synergien tussen een aantal prioritaire en secundaire terrestrische soorten
- via visuele surveys

:::


## Scenario 2: Focus on detection (plants)

- Method synergies under scopes distribution, abundance, distribution under management

```{r scenario-det-priorities-syn-plants-dist}
#| results: asis

table_syn_det_plants_display_list$distribution |> unclass() |> cat()

```


::: notes

- voor de andere scopes houden we enkel de hemelboom over
- enige prioritaire plant die wijdverspreid buiten natuurgebieden is

:::


## Scenario 2: Focus on detection (plants) 

- Again clear group: 
  - 5 terrestrial plants
  - Scope detection
  - Visual surveys
- We suggest to analyse for these plants as a group how the scope detection could be filled in



## Scenario 2: Focus on detection (animals) {.scrollable}

- Method synergies under scope detection

```{r scenario-det-priorities-syn-animals-det}
#| results: asis

table_syn_det_animals_display_list$detection |> unclass() |> cat()

```

::: notes

- realistische synergiën:
- voor een aantal aquatische soorten (primair & secundair) via eDNA
- voor een aantal terrestrische zoogdieren (primair en tertiar) via cameravallen, visuele surveys
  - vogels (primair en secundair) mogelijk erbij, maar ad hoc data vaak bruikbaar
  - andere (bv. insecten, platworm) weinig realistisch 

:::


## Scenario 2: Focus on detection (animals)

- Method synergies under scope distribution

```{r scenario-det-priorities-syn-animals-dist}
#| results: asis

table_syn_det_animals_display_list$distribution |> unclass() |> cat()

```

::: notes
- Muskusrat & blauwband via eDNA
:::

## Scenario 2: Focus on detection (animals)

- Method synergies under scope abundance

```{r scenario-det-priorities-syn-animals-abd}
#| results: asis

table_syn_det_animals_display_list$abundance |> unclass() |> cat()

```

::: notes
- geen
:::


## Scenario 2: Focus on detection (animals)

- Larger groups for scope detection
  - Group 1
    - 6 aquatic species (3 fish, Muskusrat, Californische rivierkreeft, Noord-atlantische zeester)
    - Aquatic eDNA, and for the fish electric fishing, trapping & netting
  - Group 2
    - A number of terrestrial species (mammals, birds, other)
    - Camera traps, and partly visual surveys
    - Heterogeneous group
- We suggest to analyse for the aquatic group how the scope detection could be filled in



