## Scenario 3: Focus on distribution

- Focus on species for which (also) scope distribution is relevant
  - Limited and widespread species

- Motivation
  - EU IAS regulation clearly requires reporting of distribution

::: notes
- scenario redelijk gelijkaardig aan scenario anb 
    - soorten komen redelijk sterk overeen

:::


## Scenario 3: Focus on distribution {.smaller}


::::: columns
::: {.column width="50%"}

High priority in survey


```{r scenario-dist-species}
#| results: "asis"

species_dist |> 
  dplyr::filter(grepl("highprior", scope_prior)) |>
  make_list_species_display()

```

:::
:::::


## Scenario 3: Focus on distribution (plants)

- Method synergies under scopes detection, distribution under management

```{r scenario-dist-priorities-syn-plants-det}
#| results: asis

table_syn_dist_plants_display_list$detection |> unclass() |> cat()

```

::: notes
- for scopes detection, distribution under management: Hemelboon
- geen synergien dus
:::


## Scenario 3: Focus on distribution (plants)

- Method synergies under scopes distribution, abundance 

```{r scenario-dist-priorities-syn-plants-dist}
#| results: asis

table_syn_dist_plants_display_list$distribution |> unclass() |> cat()

```

::: notes
- for scopes distribution en abundance:
- mogelijke synergie voor een groep van 5 zoetwater planten 
  - visuele surveys voor alle 5
  - eDNA: water voor 4 (mogelijk 5)
- die groep zagen we ook al onder scenario 1

:::



## Scenario 3: Focus on distribution (plants)

- Method synergies under scope abundance under management

```{r scenario-dist-priorities-syn-plants-abd-after}
#| results: asis

table_syn_dist_plants_display_list$abundance_management |> unclass() |> cat()

```

::: notes
- en voor scope abundance under management zien we die groep hier ook terug

:::


## Scenario 3: Focus on distribution (plants) 

- Clear group (cf. scenario 1):
  - 4 to 5 (partly) freshwater plants
  - Scopes distribution, abundance, and abundance under management all relevant
    - Focus on distribution most sensible
  - Aquatic eDNA, visual surveys
- We suggest to analyse for these plants as a group how the scope distribution could be filled in



## Scenario 3: Focus on distribution (animals)

- Method synergies under scope detection

```{r scenario-dist-priorities-syn-animals-det}
#| results: asis

table_syn_dist_animals_display_list$detection |> unclass() |> cat()

```


::: notes
- detectie in natuurgebieden:
- kleine groep, muskrat en blauwband

:::


## Scenario 3: Focus on distribution (animals)

- Method synergies under scope inventory

```{r scenario-dist-priorities-syn-animals-inv}
#| results: asis

table_syn_dist_animals_display_list$inventory |> unclass() |> cat()

```


::: notes
- synergie onder scope inventory zagen we ook al onder anb scenario
- blijft weinig realtistisch en de combinatie scope-soorten ook niet relevant 

::: 



## Scenario 3: Focus on distribution (animals) {.scrollable}

- Method synergies under scope distribution

```{r scenario-dist-priorities-syn-animals-dist}
#| results: asis

table_syn_dist_animals_display_list$distribution |> unclass() |> cat()

```


::: notes
- situatie heel gelijkaardig aan situatie onder anb scenario voor scope distribution:
  - synergien tussen 4 prioritaire zoetwater soorten (Blauwband is er hier bij gekomen) en gevlekte rivierkreeft als secundaire en chinese krab (ook nieuw hier) als tertiaire soort
  - via eDNA
- synergien tussen 2 terrestrische soorten en tertiaire soort wasbeer
  - cameravallen of visuele surveys

::: 



## Scenario 3: Focus on distribution (animals) {.scrollable}

- Method synergies under scope abundance

```{r scenario-dist-priorities-syn-animals-abd}
#| results: asis

table_syn_dist_animals_display_list$abundance |> unclass() |> cat()

```

::: notes
- hier ook weer gelijkenissen met anb scenario wat synergie tussen muntjak en wasbeer betreft (visuele surveys, cameravallen)
- blauwband en chinese krab mogelijk nog combineerbaar (?)

:::



## Scenario 3: Focus on distribution (animals)

- Method synergies under scope distribution under management

```{r scenario-dist-priorities-syn-animals-dist-after}
#| results: asis

table_syn_dist_animals_display_list$distribution_management |> unclass() |> cat()

```

::: notes
- exact dezelfde situatie als onder anb scenario
- synergie tussen beverrat en wasbeer weinig realistisch wegens verschillend habitat

:::



## Scenario 3: Focus on distribution (animals)

- Method synergies under scope abundance under management

```{r scenario-dist-priorities-syn-animals-abd-after}
#| results: asis

table_syn_dist_animals_display_list$abundance_management |> unclass() |> cat()

```

::: notes
- zoals in anb scenario mogelijke synergien tussen muntjak en wasbeer (visuele surveys, cameravallen)
- nieuw is de mogelijke combinatie klauwkikker & chinese krab (eDNA)oor abundance en abundance under management

:::


## Scenario 3: Focus on distribution (animals)


- Similiar groups as under scenario 1
  - Largest groups under scope distribution
    - Group 1: 3 terrestrial mammals (Aziatische grondeekhoorn, Chinese muntjak, Wasbeer) via visual surveys and camera traps
    - Group 2: 6 freshwater species (Afrikaanse klauwkikker, Beverrat, Muskusrat, Blauwband, Gevlekte rivierkreeft, Chinese krab) via eDNA
- We suggest to analyse for group 2 how the scope distribution could be filled in
