## Scenario 3: Focus on distribution {.smaller}

:::: {.columns}

::: {.column width="100%"}


- High priority species for which distribution is (also) relevant


```{r scenario-dist-species}
#| results: "asis"

species_dist |> 
  dplyr::mutate(vern_name_nld_col = kableExtra::cell_spec(
    x = vern_name_nld,
    color = dplyr::case_when(
      grepl("plant", kingdom) ~ col_plants,
      grepl("dier", kingdom) ~ col_animals
    )
  )) |>
  dplyr::arrange(vern_name_nld) |>
  dplyr::pull(vern_name_nld_col) |>
  paste(x = "  -", y = _) |>
  cat(x = _, sep = "\n")

```


:::

::::



## Scenario 3: Focus on distribution (plants)

- Method synergies under scopes detection, distribution after management

```{r scenario-dist-priorities-syn-plants-det}
#| results: asis

table_syn_dist_plants_display_list$detection |> unclass() |> cat()

```


## Scenario 3: Focus on distribution (plants)

- Method synergies under scopes distribution, abundance 

```{r scenario-dist-priorities-syn-plants-dist}
#| results: asis

table_syn_dist_plants_display_list$distribution |> unclass() |> cat()

```



## Scenario 3: Focus on distribution (plants)

- Method synergies under scope abundance after management

```{r scenario-dist-priorities-syn-plants-abd-after}
#| results: asis

table_syn_dist_plants_display_list$abundance_management |> unclass() |> cat()

```

## Scenario 3: Focus on distribution (plants) 

- Cf. scenario 1: 
  - 4 to 5 (partly) freshwater plants
  - Scopes distribution, abundance, and abundance after management
  - Aquatic eDNA, visual surveys
- We suggest to analyse for these plants as a group which scopes can be filled in how, at which costs


::: {.notes}

nota's: 

- hier graag jullie input

:::



## Scenario 3: Focus on distribution (animals)

- Method synergies under scope detection

```{r scenario-dist-priorities-syn-animals-det}
#| results: asis

table_syn_dist_animals_display_list$detection |> unclass() |> cat()

```

## Scenario 3: Focus on distribution (animals)

- Method synergies under scope inventory

```{r scenario-dist-priorities-syn-animals-inv}
#| results: asis

table_syn_dist_animals_display_list$inventory |> unclass() |> cat()

```



## Scenario 3: Focus on distribution (animals)

- Method synergies under scope distribution

```{r scenario-dist-priorities-syn-animals-dist}
#| results: asis

table_syn_dist_animals_display_list$distribution |> unclass() |> cat()

```

## Scenario 3: Focus on distribution (animals)

- Method synergies under scope abundance

```{r scenario-dist-priorities-syn-animals-abd}
#| results: asis

table_syn_dist_animals_display_list$abundance |> unclass() |> cat()

```



## Scenario 3: Focus on distribution (animals)

- Method synergies under scope distribution after management

```{r scenario-dist-priorities-syn-animals-dist-after}
#| results: asis

table_syn_dist_animals_display_list$distribution_management |> unclass() |> cat()

```



## Scenario 3: Focus on distribution (animals)

- Method synergies under scope abundance after management

```{r scenario-dist-priorities-syn-animals-abd-after}
#| results: asis

table_syn_dist_animals_display_list$abundance_management |> unclass() |> cat()

```

## Scenario 3: Focus on distribution (animals)

- Similiar groups as under scenario 1
  - Scope distribution
    - Group 1 (scenario 1)
    - Group 2 (scenario 1 & Stone moroko) 
  - Scopes abundance
    - Chinese muntjac & African clawed (scenario 1) & Stone moroko
    - Unique method synergies for African clawed & Stone moroko  
- Possibility to group African clawed & Stone moroko in addition to suggestions under scenario 1 (?)


::: {.notes}

nota's: 

- hier graag jullie input

:::
