#  

## Scoring questions

Distribution and abundance


```{r survey-questions-scoring-section-3}

make_scoring_table(
  data_table_scoring_upd  |>
  dplyr::filter(section_number == 3) |>
  dplyr::select(-c("section_number", "score_category")) 
) |> gsub(
  pattern = "feasibility",
  replacement = kableExtra::cell_spec(x = "feasibility", color = col_feas),
  x = _
) |> gsub(
  pattern = "urgency",
  replacement = kableExtra::cell_spec(x = "urgency", color = col_urge),
  x = _
)

```



## Scoring questions

Impact


```{r survey-questions-scoring-section-4}

make_scoring_table(
  data_table_scoring_upd  |>
  dplyr::filter(section_number == 4) |>
  dplyr::select(-c("section_number", "score_category")) 
) |> gsub(
  pattern = "feasibility",
  replacement = kableExtra::cell_spec(x = "feasibility", color = col_feas),
  x = _
) |> gsub(
  pattern = "urgency",
  replacement = kableExtra::cell_spec(x = "urgency", color = col_urge),
  x = _
)

```


## Scoring questions

Monitoring 


```{r survey-questions-scoring-section-5}

make_scoring_table(
  data_table_scoring_upd  |>
  dplyr::filter(section_number == 5) |>
  dplyr::select(-c("section_number", "score_category")) 
) |> gsub(
  pattern = "feasibility",
  replacement = kableExtra::cell_spec(x = "feasibility", color = col_feas),
  x = _
) |> gsub(
  pattern = "urgency",
  replacement = kableExtra::cell_spec(x = "urgency", color = col_urge),
  x = _
)

```

## Scoring questions

Management


```{r survey-questions-scoring-section-6}

make_scoring_table(
  data_table_scoring_upd  |>
  dplyr::filter(section_number == 6) |>
  dplyr::select(-c("section_number", "score_category")) 
) |> gsub(
  pattern = "feasibility",
  replacement = kableExtra::cell_spec(x = "feasibility", color = col_feas),
  x = _
) |> gsub(
  pattern = "urgency",
  replacement = kableExtra::cell_spec(x = "urgency", color = col_urge),
  x = _
)

```




## Scope distribution (EU)


:::: {.columns}

::: {.column width="50%"}

![Gap closure algorithm (EU guidelines)](media/EU_habitatdir_guidelines_fig4.svg){height=400}

:::

::: {.column width="50%"}

![Occurrence, distribution, range (EU guidelines)](media/EU_habitatdir_guidelines_fig3.svg){height=400}

:::

::::


::: {.notes}
- hoe wordt de range bepaald?
- daarvoor gebruik gemaakt van een spatiale algorithme, met naam 'gap closure'
- 'gap distance'
- range continuity: distance between 2 distribution grids < 'gap distance'
- range discontinuity: distance between 2 distribution grids > 'gap distance'
- tool beschikbaar
- deduced range is in a second step adapted using information on unsuitable areas 
  - e.g., marine areas in range of terrestrial species
  - e.g., areas without freshwater bodies for freshwater species
- relationship occurrence, distribution, range

:::



## Scope distribution (EU)


![Species-specific 'gap distances' (EU guidelines)](media/EU_habitatdir_guidelines_tab7.svg){height=300}

- Optimize study design for given 'gap distance'

::: {.notes}
- 'gap distances' are species-specific and should reflect ecological characteristics of the species (e.g., mobility)
- 'gap distances' belangrijk omdat die bepaald hoe een geoptimaliseerd design eruit zou zien,
    - ruimtelijke resolutie die nodig is om data te verzamelen die de nodige informatie bevat
    - hokken daartussen / hokken in hogere resolutie hebben geen impact op range en kunnen bv. met ad-hoc data ingevuld worden

:::


## Data quality

![Types of data collection (De Knijf et al., 2014)](media/dataquality.svg){height=500}



